---
title: "homework 2"
author: "Sze Pui Tsang"
date: "9/10/2021"
output: github_document
---
load the library first
```{r library}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
```

# Problem 1

```{r define path}
path_of_data = "C:/Users/Yang/Desktop/st3340_p8105_classex/data_hw2/Trash-Wheel-Collection-Totals-7-2020-2.xlsx"
```

```{r import the dataframe}
mr_trash_df = 
  read_excel( 
    path = path_of_data,
    sheet = "Mr. Trash Wheel"
    )
```
We have to clean the table by cleaning up the variable names,
omitting the missing value abd rounding the number of sport balls to 
the nearest integer

```{r clean the data frame}
clean_mr_trash_df = mr_trash_df %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate ( 
    sports_balls = round(sports_balls, digits = 0)
  )
  
``` 

Read and clean precipitation data for 2018. 
```{r Read data for 2018}
precipitation_18_df= read_excel( 
    path = path_of_data,
    sheet = "2018 Precipitation",
    range = "A2:B14")
```

```{r clean data for 2018}
precipitation_18_df = precipitation_18_df %>%
  janitor :: clean_names() %>%
  drop_na(month) %>%
  mutate(year =2018) %>%
  relocate(year, .before = "month")%>%
  mutate(
    total = round(total, digit = 1)
  )
```

Read and clean precipitation data for 2019.
```{r clean data for 2019}
precipitation_19_df= read_excel( 
    path = path_of_data,
    sheet = "2019 Precipitation",
    range = "A2:B14")%>%
  janitor :: clean_names() %>%
  drop_na(month)%>%
  mutate(year =2019)%>%
  relocate(year, .before = "month")%>%
  mutate(
    total = round(total, digit = 1)
  )
```

Combind precipitation dataset and convert month into a character variable
```{r combind the two dataset}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name)
precipitation_18and19_df = bind_rows(precipitation_18_df,precipitation_19_df)
precipitation_18and19_df = left_join(precipitation_18and19_df, month_df, by="month")
```

view the dataset
```{r}
median(precipitation_19_df$total)
sum(precipitation_18_df$total)
```

The median number of sports balls in a dumpster in 2019 was 3.35
The total precipitation in 2018 was 70.3 inches


# Problem 2
```{r import the dataframe}
pols_month_df = read.csv("pols-month.csv", header = TRUE, sep= ",")
snp_df =read.csv("snp.csv", header = TRUE, sep= ",")
wide_umemployment_df = read.csv("unemployment.csv", header = TRUE, sep= ",")
```

Firstly, clean the data in pols- months
learning note: R ifelse function 
ifelse(testexpression, x, y)
The returned vector has element from x if the corresponding value of test_expression 
is TRUE and return y if it is FALSE 

```{r clean dataframe for pol_month}
clean_pols_month_df = pols_month_df %>%
  separate(mon, c("year", "month", "day"), sep = "-") %>%
  mutate (month = month.name[as.integer(month)])%>%
  mutate(year = as.integer(year))%>%
  mutate (president = ifelse(prez_dem == 1, "dem", "gop"))%>%
  select(-prez_dem, -prez_gop, - day)%>%
  arrange(year,month)
```

Clean the np dataset
It is noticed that the column year has ony two digits, 
ifelse function is applied in order to turn them into four digits, 
which is more readable 

```{r clean dataframe for snp}
clean_snp_df = snp_df %>%
  separate(date, c("day", "month", "year"), sep = "/") %>%
   mutate (month = month.name[as.integer(month)])%>%
  mutate(year = as.integer(year))%>%
  relocate(year, month)%>%
  arrange(year,month)%>%
  mutate(year = ifelse( year<15, year+2000, year+1900))%>%
  select(-day)
```

Joining the two dataset snp and pols together 

```{r}
snp_pols_df = left_join(clean_snp_df, clean_pols_month_df, by=c("year","month"))
```

Turning unemplyment from long to wide 
```{r}
long_umemployment_df = wide_umemployment_df%>%
  janitor::clean_names()%>%
  pivot_longer(
    cols=!year,
    names_to="month", 
    values_to = "unemployment rate")%>%
    mutate(year = as.integer(year))
```

Merge the three tables together
```{r}

umeployment_snp_pols = bind_rows(snp_pols_df,long_umemployment_df) 
```
#Problem 3
```{r import the dataframe}
baby_name_df = read_csv("Popular_Baby_Names.csv", show_col_types = FALSE)
view(baby_name_df)
```

```{r clean the data}
baby_name_df = baby_name_df%>%
  janitor :: clean_names()%>%
  distinct()%>%
  mutate(across(where(is.character), toupper))%>%
  mutate(ethnicity = replace(ethnicity,ethnicity =="BLACK NON HISP", "BLACK NON HISPANIC"),
         ethnicity = replace(ethnicity,ethnicity =="WHITE NON HISP", "WHITE NON HISPANIC"),
         ethnicity = replace(ethnicity,ethnicity =="ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"))

```



```{r a dataframe ranking the popularity of female's baby name "Olivia}

Olivia_rank_df = baby_name_df%>%
  filter(childs_first_name == "OLIVIA")%>%
  select(-gender,-childs_first_name, -count)%>%
  pivot_wider(names_from ="year_of_birth", values_from = "rank")
```


```{r a dataframe ranking the most popular male chidren over time}
popular_male_name = baby_name_df %>%
  filter(gender == "MALE", rank == 1)%>%
  select(-gender, -count, -rank)%>%
  pivot_wider(names_from ="year_of_birth", values_from = "childs_first_name")
```

```{r scatter plot}
scatter_plot = baby_name_df %>%
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) 

scatter_plot_2 = ggplot(scatter_plot, aes(x = rank, y = count)) + geom_point()
print(scatter_plot_2)
```
